Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalogue Data

### 2012-12-04 doc/paper-lcr.tex
draft update
acted on phils comments

\documentclass[useAMS,usenatbib]{mn2e}
%% letterpaper
%% a4paper

\voffset=-0.6in

@@

This procedure of reducing the full PDF for $\kappah$ to its median increases
the importance of the realism of the simulation: the simulation needs to be a
similar as possible to the real universe or there is a possibility of
systematic errors. \phil{On the other hand, making use of all the information
in the photometric catalogue yields $\kappax$ estimates of increased
precision, as we shall see.}

@@

\Fref{fig:jointkh-k} shows the conditional distribution of $\kappax$ given 
$\kappah$, derived from $10^5$ randomly selected \MS lines of sight
and assuming noise-free halo masses and redshifts. 
\phil{(In this case, $\kappah^{\rm med} = \kappah$, since 
$\pr(\kappah|\data)$ is a delta function)}.
% In this case, the PDF $\pr(\kappah|\data)$ is a delta function: we can read
% off  $\pr(\kappax|\data)$ just by taking a thin vertical slice through this 
% distribution. 
% Note that the joint PDF plotted in this figure is the product
% of  the conditional distribution that we need for calibration, and also the
% prior PDF for $\kappah$ that we need to ensure that the global distribution of
% $\kappah$ values is defaulted to in the case of poor observational data. 
At fixed $\kappah$ we find in \Fref{fig:jointkh-k} that the scatter in 
$\kappax$
grows with $\kappah$; our reconstruction is better at reproducing under-dense
lines of sight than over-dense lines of sight. The effect of ignoring voids
and the smooth mass component is evident in this plot: $\kappah$ is
significantly higher than $\kappax$, at any given $\kappax$ value. 
\phil{Conditioning on $\kappah$ gives a PDF for
$\kappax$ centred at the correct place, by construction.}

@@

\flag{Should we show the relevant calibration plots, analogous to Figure 2?
Seems a bit odd to give Figure 2 but not illustrate the {\it actual
calibration procedure we follow on real data}... Maybe we could make one that
had the conditional distribution for each data quality?}

@@

\comment{PJM: The following does not belong here because \Fref{fig:widthsvsH}
does not show results for spectroscopic redshifts only!

Reconstructing the lines of sight with perfect knowledge of the
redshift but an uncertain stellar mass, we find that the width of
$\pr(\kappax|\data)$ grows with the expectation value of $\kappax$
(\Fref{fig:widthsvsH}). The low $\kappax$ lines of sight are relatively empty,
and so there are few opportunities for uncertainties in the halo masses to
\propogate into $\kappax$ uncertainties. 
The median reconstruction width in
this idealised situations is 0.01; some 22\% of lines of sight would have a
reconstruction width less than 0.01. }

% PJM: Note that I removed talk of void correction here - surely we should be
% just applying the method we describe and inferring $\kappa$?
%
% TEC: yes.

@@

@@

\flag{Is there a simple approximation for $\kappax$ and $\sigma_{\kappa}$ as a
function of $\kappaxtrue$? This would be useful when making forecasts, writing
proposals etc...}

@@

While a fully photometric reconstruction provides useful constraints on
$\pr(\kappax|\data)$, targeted spectroscopy can provide additional constraints
on the masses and redshifts of halos whose $\kappah$ contributions have the
largest absolute uncertainty. Obtaining spectra of bright ($i<22$) objects is
relatively fast with large telescopes: however, we find that if spectroscopic
redshifts were known for all $i<22$ galaxies in our fields the reconstruction
improves only slightly over the purely photometric reconstruction,  although
the improvement can depend strongly on the details of the particular line of
sight. Obtaining spectra for fainter objects would be correspondingly more
expensive, but if spectroscopic redshifts could be obtained for all $i<23$
galaxies (as part of a futuristic baryon acoustic oscillation survey, for
example) and all $i<24$ galaxies within 1 arcminute of the line of sight, then
we find that the $\pr(\kappax|\data)$ would be almost as precise as that from
having complete spectroscopic coverage (\Fref{fig:reconwidths}). High
$\kappax$ lines of sight benefit the most from additional spectral coverage,
since these have the most uncertain $\kappah$. \flag{This is not what
\Fref{fig:widthsvsH} shows at all! Does the $\kappaxtrue$ axis not extend far
enough?} Consistent with our previous findings, the lowest $\kappax$ lines of
sight have the most constrained $\kappax$ PDFs.

@@

While it is good to reconstruct $\kappax$ precisely, it is more important that
$\kappax$ is reconstructed accurately. The reconstruction must not be biased. 
We quantify the accuracy of our reconstructed $\pr(\kappax|\data)$ in the
following way. By shifting the \infered $\pr(\kappax|\data)$ for each line of
sight by the true (ray-traced) convergence, we obtain a  PDF that should be
centred on zero; these offset PDFs can be multiplied together for multiple
lines of sight to emulate a joint likelihood analysis, and then test for
possible bias in it:
\be
\label{eq:bias}
\mathcal{P}_N = \prod_{i=1}^N \pr_i(\kappax-\kappax^{\rm true}|\data)
\ee
In this section, we quantify the bias as the size of the deviation in the
expectation value of $\mathcal{P}_N$ from zero. If the bias of $\mathcal{P}_N$
is smaller than the half-width of $\mathcal{P}_N$, our $\kappax$ inferences
can be considered accurate. Since the width of $\mathcal{P}_N$ decreases as
more sightlines are combined, a small bias will always 
eventually be found. In the context of measuring time delay distance at the 
one percent level, the expectation value of $|\mathcal{P}_N|$ needs to be 
significantly less than 0.01.
\flag{Modify this as advised by Sherry -- introduce model for other
uncertainties in lenses, and explain number of lenses to see error argument.}

@@

We first consider a purely random selection of lines of sight. With the purely
photometric reconstruction, there is no evidence of bias, even after combining
20 lines of sight. \phil{This result is really a sanity check, since for
random lines of sight the bias as defined here should be zero by
construction}. If these lines of sight hosted time delay lenses, the bias
number would correspond to a systematic error of 0.5\% in time delay distance.
\flag{What is this number? In the first sentence the bias was given as zero!}
Since the statistical distance uncertainties due to lens modelling and time
delay estimation are likely to be at the 3-4\% level \citep{SuyuEtal2010}, 
only a sample containing 200 or more lenses would be limited by the residual
bias following external convergence reconstruction.
\flag{$0.035/\sqrt{200} \neq 0.005$ so what numbers are these? We should
outline the lens number calculation in the preamble above.}

@@

We find that selecting the most tightly constrained lines of sight in this way
does not introduce a bias: 
our reconstruction is most accurate for the lowest
$\kappax$ lines of sight. With targeted spectroscopic follow-up of the
selected lines of sight there is only a slight 
improvement compared to photometry
alone, since the best constrained lines of sight are also 
the most empty, and it is
the high-$\kappax$ lines of sight that benefit the most from spectroscopy.
\flag{This was not shown conclusiveley earlier. Need to tighten this up.}
Photometry of the field seems to be sufficient to adequately select these
empty lines of sight.  However, 
since lenses are typically found in locally over-dense
environments, it is unlikely that  an almost empty line of sight would
actually  contain a lens; \phil{what we can say is that if the local
environment can be identified spectroscopically, and then the line
of sight mass distribution reconstructed using our method, selecting lenses
with low, well-constrained values of $\kappax$ will lead to an increase in
distance precision at no cost in bias.}

@@

\flag{Paragraph outlining future work on lens lines of sight (eg including
local environment). Why wasn't this done here?}

@@

@@

\item Selecting from the highest precision 33\% of the reconstructed lines of
sight results in a sample with low bias, 0.1\% in terms of the time delay
distance, and a statistical uncertainty due to line of sight mass structure
that is half that of the parent sample. \flag{This is the first time this
number (0.1\%) has appeared!} Neglecting other sources of uncertainty, the
residual systematic error would be equal to the overall statistical
uncertainty in a sample of around 200 lenses. \flag{Needs checking.}

>>>>

\documentclass[useAMS,usenatbib,a4paper]{mn2e}
%% letterpaper
%% a4paper

%\voffset=-0.6in

@@

This procedure of reducing the full PDF for $\kappah$ to its median increases
the importance of the realism of the simulation: the simulation needs to be a
similar as possible to the real universe or there is a possibility of
systematic errors. Conversely, making use of all the information
in the photometric catalogue yields $\kappax$ estimates of increased
precision.

@@

\Fref{fig:jointkh-k} shows the conditional distribution of $\kappax$ given 
$\kappah$, derived from $10^5$ randomly selected \MS lines of sight
and assuming noise-free halo masses and redshifts. 
\phil{(In this case, $\pr(\kappah|\data)$ is almost delta function since the only uncertainty is halo
concentration, which has minimal effect on $\pr(\kappah|\data)$.  
)}.
% In this case, the PDF : we can read
% off  $\pr(\kappax|\data)$ just by taking a thin vertical slice through this 
% distribution. 
% Note that the joint PDF plotted in this figure is the product
% of  the conditional distribution that we need for calibration, and also the
% prior PDF for $\kappah$ that we need to ensure that the global distribution of
% $\kappah$ values is defaulted to in the case of poor observational data. 
At fixed $\kappah$ we find in \Fref{fig:jointkh-k} that the scatter in 
$\kappax$
grows with $\kappah$; our reconstruction is better at reproducing under-dense
lines of sight than over-dense lines of sight. The effect of ignoring voids
and the smooth mass component is evident in this plot: $\kappah$ is
significantly higher than $\kappax$, at any given $\kappax$ value. 
\phil{Conditioning on $\kappah$ gives a PDF for
$\kappax$ centred at the correct place, by construction.}

@@

%\flag{Should we show the relevant calibration plots, analogous to Figure 2?
%Seems a bit odd to give Figure 2 but not illustrate the {\it actual
%calibration procedure we follow on real data}... Maybe we could make one that
%had the conditional distribution for each data quality?}

%TEC: I think that could lead to figure overload; the concept is given throughout
% the paper, and it's well illustrated by figure 2.

@@

\comment{PJM: The following does not belong here because \Fref{fig:widthsvsH}
does not show results for spectroscopic redshifts only!}

Reconstructing the lines of sight with perfect knowledge of the
redshift but an uncertain stellar mass, we find that the width of
$\pr(\kappax|\data)$ grows with the expectation of $\kappax$ and the ray-traced $\kappatrue$
(\Fref{fig:widthsvsH}). The low $\kappax$ lines of sight are relatively empty,
and so there are few opportunities for uncertainties in the halo masses to
\propogate into $\kappax$ uncertainties; low $\kappax$ lines of sight can be reconstructed
more precisely than high $\kappax$ lines of sight.

@@

\tom{With photometric data alone, $\kappah^{\rm med}$ can shift by $\sim$0.01 relative to $\kappah^{\rm med}$
given spectroscopic coverage. The $\kappax$ contribution of voids cannot change with data quality;
the shift in $\kappah^{\rm med}$ is due to the asymmetric propogation of stellar mass uncertainties
into $\kappax$ uncertainties. As stellar mass uncertainty increases $\kappah^{\rm med}$ is pushed higher. 
To correctly calibrate  $\pr(\kappax|\kappah^{\rm med},\data)$ one {\it must} include the data quality used to generate $\kappah^{\rm med}$.}

@@

%\flag{Is there a simple approximation for $\kappax$ and $\sigma_{\kappa}$ as a
%function of $\kappaxtrue$? This would be useful when making forecasts, writing
%proposals etc...}
%
%TEC: No. It's also not relevant to proposals since kappatrue isnt observable. One can ask is there
%a formula for sigmakappa given the expectation value of kappax. The answer to that is
%yes, but it depends on data quality. Concerning forecasting yes we would 
%want to use a simple approximation, but I don't think it has a place in this paper.
%When you/I/someone does the forecasting I'd like to be involved; I'll make the
%approximation then.

@@

While a fully photometric reconstruction provides useful constraints on
$\pr(\kappax|\data)$, targeted spectroscopy can provide additional constraints
on the masses and redshifts of halos whose $\kappah$ contributions have the
largest absolute uncertainty. Obtaining spectra of bright ($i<22$) objects is
relatively fast with large telescopes: however, we find that if spectroscopic
redshifts were known for all $i<22$ galaxies in our fields the reconstruction
improves only slightly over the purely photometric reconstruction,  although
the improvement can depend strongly on the details of the particular line of
sight. Obtaining spectra for fainter objects would be correspondingly more
expensive, but if spectroscopic redshifts could be obtained for all $i<23$
galaxies (as part of a futuristic baryon acoustic oscillation survey, for
example) and all $i<24$ galaxies within 1 arcminute of the line of sight, then
we find that the $\pr(\kappax|\data)$ would be almost as precise as that from
having complete spectroscopic coverage (\Fref{fig:reconwidths}).
Consistent with our previous findings, the lowest $\kappax$ lines of
sight have the most constrained $\kappax$ PDFs.

@@

While it is good to reconstruct $\kappax$ precisely, it is more important that
$\kappax$ is reconstructed accurately. The reconstruction must not be biased. 
We quantify the accuracy of our reconstructed $\pr(\kappax|\data)$ in the
following way. By shifting the \infered $\pr(\kappax|\data)$ for each line of
sight by the true (ray-traced) convergence, we obtain a  PDF that should be
centred on zero; these offset PDFs can be multiplied together for multiple
lines of sight to emulate a joint likelihood analysis, and then test for
possible bias in it:
\be
\label{eq:bias}
\mathcal{P}_N = \prod_{i=1}^N \pr_i(\kappax-\kappax^{\rm true}|\data)
\ee
In this section, we quantify the bias as the size of the deviation in the
expectation value of $\mathcal{P}_N$ from zero. If the bias of $\mathcal{P}_N$
is smaller than the half-width of $\mathcal{P}_N$, our $\kappax$ inferences
can be considered accurate. Since the width of $\mathcal{P}_N$ decreases as
more sightlines are combined, a small bias will always 
eventually be found. In the context of measuring time-delay distance at the 
one percent level, the expectation value of $|\mathcal{P}_N|$ needs to be 
significantly less than 0.01. \tom{External convergence is not the only
source of uncertainty in measuring time-delay distance.
Since the statistical distance uncertainties due to lens modelling and time
delay estimation are likely to be at the 3-4\% level \citep{SuyuEtal2010},
convolving $\mathcal{P}_N$ with a Gaussian of width $0.03/\sqrt{N}$ is a
reasonable approximation for the 
likely final uncertainty on time-delay distance.}

@@

We first consider a purely random selection of lines of sight. With the purely
photometric reconstruction, there is no evidence of bias, even after combining
20 lines of sight. \tom{This result is a sanity check: if we had chosen lines of
sight with identical $\kappah^{\rm med}$ rather than $\kappah^{\rm med}\pm0.003$ randomly
selected lines of sight would have zero bias by construction}. If lines of sight hosting time delay lenses
are truely random,  our method allows $\kappax$ to be reconstructed without inducing a bias\footnote{under the assumption that the universe is like the calibration simulation, and the correct stellar-mass to halo-mass relation has been used}.

@@

We find that selecting the most tightly constrained lines of sight in this way
does not introduce a bias: 
our reconstruction is most accurate for the lowest
$\kappax$ lines of sight. Combining 20 such systems the bias is at the 0.001 level.

Photometry of the field seems to be sufficient to adequately select these
empty lines of sight.  However, 
since lenses are typically found in locally over-dense
environments, it is unlikely that  an almost empty line of sight would
actually  contain a lens; \phil{what we can say is that if the local
environment can be identified spectroscopically, and then the line
of sight mass distribution reconstructed using our method, selecting lenses
with low, well-constrained values of $\kappax$ will lead to an increase in
distance precision at no cost in bias.}

@@

@@

\tom{ Future work should focus on two areas:
reducing the dependance on simulations; and,
improving the modeling of dark structures and voids. The analysis should be repeated with different 
simulations and differrent stellar-mass to halo-mass relations, cross-checking that 
using the wrong simulation simulation or relation does not induce a systematic error.
Improving the modelling of of voids could be done using the void models of \citet{voids2}
or \citet{voids1} combined with a void finder such as in \citet{voids3}.
Combining the halo model with a void model may allow a direct test of the 
halo approximation; using real photometric data one could reconstruct lines of sight
and compare against measured weak lensing shears. This would ensure that the halo model reproduces the
contents of the real universe, rather than a particular simulation.}

@@

\item Selecting from the highest precision 33\% of the reconstructed lines of
sight results in a sample with low bias, $\sim$0.1\% in terms of the time-delay
distance. \tom{For these lines of sight the reconstructed $\pr(\kappax)$ induce 
a statistical uncertainty of $\sim$1\% on individual time-delay distances}
